Using Machine Learning Tools for Open Cluster Characterization with Gaia DR2 Data

### 2020-12-10 thesis/thesis.tex

\section{Deep Embedded Clustering (DEC)}

@@

For that reason we have adapted the \emph{Unsupervised Deep Embedding for Clustering Analysis} or \emph{DEC} model \cite{xie2016unsupervised} to our work.

@@

@@

% Se aporta la discusión de los resultados presentados en el capítulo anterior. En este capítulo se puede discutir la relevancia de
% los resultados, presentar posibles explicaciones para los datos anómalos y resaltar aquellos datos que sean particularmente relevantes
% para el análisis del experimento.

@@

% Resumen final del trabajo y debe servir para informar del alcance y relevancia de la aportación.
% Suele estructurarse empezando con un resumen del problema tratado, de cómo se ha abordado y de por qué la solución sería válida.

% Es recomendable que incluya también un resumen de las contribuciones del trabajo, en el que se relacionen las contribuciones y los
% resultado obtenidos con ls objetivos planteados para el trabajo, discutiendo hasta qué punto has conseguido resolver los objetivos planteados.

@@

% Debería señalar las perspectivas de futuro que abre el trabajo desarrollado para el campo de estudio definido. En el fondo,
% debe justificar de qué modo puede emplearse la aportación desarrollada y en qué campos.

@@

\appendix
\chapter{Feature Selection}
\label{chap:feature_selection}

>>>>

\section{Deep Embedded Clustering (DEC)}
\label{sec:deep_embedding_clustering}

@@

For that reason we have adapted the \emph{Unsupervised Deep Embedding for Clustering Analysis} to our work.

@@

All results have been computed with an Apple Mac Pro Late 2013, 2.7 GHz 12-Core Intel Xeon E5 with two AMD FirePro D700 6GB and 64 GB
of memory, running macOS BigSur 11.0.1, Keras 2.2 and PlaidML 0.7.

The majority of the runnings have taken a few minutes, while big clusters such as Melotte 25, have taken around an hour to perform
the computations. This contrasts with \citeA{castro2020hunting} work, that used the power of the MareNostrum 42 supercomputer in order
to perform its calculations.

@@

We have applied the our proposed method to clusters with different topologies:

\begin{itemize}
  \item NGC 2516 has it proper motion center deviated from the origin but it is embedded inside a big cloud of stars with similar proper motions.
  \item NGC 2632 (in addition to Melotte 22) ia s cluster whose proper motion center is not located at the origin and has a well separated parallax center.
  \item NGC 2682, on the other hand, has its parallax centered within the region's gaussian,
        which complicates its detection although its proper motion center is deviated from the origin.
  \item Melotte 25 (as well as Melotte 22) is a cluster closer to us than the other.
        That makes its memebership stars to be more scattered than previous clusters which are more compact.
\end{itemize}

All of these clusters have a wide variety of diameters, from 25 to 330 arcmin, as well as the number of stars that belong to them.
NGC 2682 has 3,000 stars while Melotte 25 is composed by more than 400,000 stars.

% TODO: Meter una tabla con los parámetros de los clusters mencionados

% TODO: Determinar correctamente los errores en las medidas y mencionarlos

In all cases, the model has resolved properly the identity of the cluster and has characterized the membership stars,
giving results compatible with the ones obtained with classic procedures and VO tools.

Compared with the mentioned tools, our model can be fully calified as a non-parameterized model since it does not depend on parameters referred
to the clusters itself but only hyperparameters such as the initial number of clusters to find, or the structure of the Clustering Layer used by DEC model.

Our model is also non-supervised since we do not need to tell the model which stars belong to the cluster or assit it while is training.

Only at the end of the characterization, a fine tunning driven by the user, can be applied in orther to improve the selection by discarding those
stars which fall outside the given quantiles.

In any case we make assumptions about the cluster profile, something that we have to take into account when using VO tools.
This is another evidence that our model is non-parameterized.

@@

As shown by the obtained results, we can say that in general our model works fine identifying and characterizing open clusters.
So we have succedded building a model non-parameterized non-supervised for open cluster characterization.
However, some improvements should be done in order to improve its accuracy and precision.

We have built a non-parameterized model since the user does not provide any information about the cluster profile or its properties.
Although some tweaks can be done regarding the cluster, like changing the initial number of clusters or modifying the Clustering Layer structure,
in order to improve the final characterization.

Another key point in our model is the initializer kernel.
This kernel prepares data before passing it to the ANN and the results may vary significantly depending on this kernel.
For example, simply varying the kernel seed can make a huge difference on the result.

Even so, comparing our method with Clusterix, we can still say that the number of parameters (or hyperparameters) in our model is smaller
and we do not need previous knowledge about the studied cluster.
That makes easier to test different configurations and automate the process.

With all of this points in mind, we can affirm that this model could be really considered as an alternative model for open cluster characterization and thus,
it could even be incloud as part of the Virtual Observatory toolset.

@@

The first aim after finishing this work is testing the model with a wider range of clusters.
In general we could think of applying it to the whole VizieR catalogue.
That way we could have a better idea about the current limits of our model and we could make adjustments to improve it.
We could determine with better accuracy the topologies that our model works well with, and which ones out model does not resolve properly.
Or even, we could stablish different sets of hyperparameters regarding the topology of the cluster.

With our model, new secondary clusters arise from the data apart of the main ones.
So, it is also interesting the individual study of these new clusters.

Another possible change in our model would be using DBSCAN instead of KMeans as our initial clustering algorithm.
Maybe this algorithm gives us a better starting point for the DEC model.

At last, we have used during this work the Gaia DR2 database, but recently the DR3 dataset has been released.
So it is evident the interest on testing our model with this new data.

@@

\appendix
